The Chemistry Development Kit (CDK): atom typing, rendering, molecular formulas, and substructure searching

### 2016-10-10 article.tex
Per request from Kai
Added a citation to CSI:FingerID

\author[
   addressref={um},                                 % ORCID: 0000-0001-7542-0286
   %corref={aff1},
   %noteref={n1},
   email={egon.willighagen@maastrichtuniversity.nl}
]{\inits{ELW}\fnm{Egon L} \snm{Willighagen}}
\author[
   addressref={nm},
   email={john.may@cantab.net}
]{\inits{JWM}\fnm{John W} \snm{May}}
\author[
   addressref={uppsala},
   email={jonathan.alvarsson@farmbio.uu.se}
]{\inits{JA}\fnm{Jonathan} \snm{Alvarsson}}
\author[
   addressref={uppsala},
   email={berg.arvid@gmail.com}
]{\inits{AB}\fnm{Arvid} \snm{Berg}}
\author[
   addressref={azg},                                % ORCID: 0000-0001-9491-4134
   email={Lars.A.Carlsson@astrazeneca.com}
]{\inits{LC}\fnm{Lars}~\snm{Carlsson}}
\author[
   addressref={jena},
   email={kai.duehrkop@uni-jena.de}
]{\inits{KD}\fnm{Kai} \snm{Dührkop}}
\author[
   addressref={idea},
   email={jeliazkova.nina@gmail.com}
]{\inits{NJ}\fnm{Nina} \snm{Jeliazkova}}
\author[
   addressref={leicester},
   email={shk12@le.ac.uk}
]{\inits{SK}\fnm{Stefan} \snm{Kuhn}}
\author[
   addressref={wi_mit},                             % ORCID: 0000-0002-6940-3006
   email={pluskal@wi.mit.edu}
]{\inits{TP}\fnm{Tomáš}~\snm{Pluskal}}
\author[
   addressref={miquel},                             % ORCID: 0000-0002-4659-1446
   email={mrojas@qca.es}
]{\inits{MRJ}\fnm{Miquel}~\snm{Rojas-Chertó}}
\author[
   addressref={uppsala},
   email={ola.spjuth@farmbio.uu.se}
]{\inits{OS}\fnm{Ola} \snm{Spjuth}}
\author[
   addressref={gilleain},                           % ORCID: 0000-0002-8368-6954
   email={gilleain.torrance@gmail.com}
]{\inits{GT}\fnm{Gilleain} \snm{Torrance}}
\author[
   addressref={um},
   email={chris.evelo@maastrichtuniversity.nl}
]{\inits{CTE}\fnm{Chris T} \snm{Evelo}}
\author[
   addressref={nih},
   email={guhar@mail.nih.gov}
]{\inits{RG}\fnm{Rajarshi} \snm{Guha}}
\author[
   addressref={ebi},
   email={steinbeck@ebi.ac.uk}
]{\inits{CS}\fnm{Christoph}~\snm{Steinbeck}}

\address[id=um]{
  \orgname{Dept of Bioinformatics - BiGCaT, NUTRIM, Maastricht University}, % university, etc
  %\street{Waterloo Road},
  \postcode{NL-6200 MD},
  \city{Maastricht},
  \cny{The Netherlands}
}
\address[id=nm]{
  \orgname{NextMove Software Ltd},
  %\street{D\"{u}sternbrooker Weg 20},
  \postcode{CB4 0EY},
  \city{Cambridge},
  \cny{UK}
}
\address[id=uppsala]{
  \orgname{Department of Pharmaceutical Biosciences, Uppsala University},
  %\street{D\"{u}sternbrooker Weg 20},
  \postcode{751 24},
  \city{Uppsala},
  \cny{Sweden}
}
\address[id=azg]{
  \orgname{AstraZeneca, Innovative Medicines \& Early Development, Quantitative Biology},
  %\street{},
  %\postcode{LE1 7RH},
  \city{Mölndal},
  \cny{SE}
}
\address[id=jena]{
  \orgname{Chair for Bioinformatics, Friedrich Schiller University},
  %\street{},
  \postcode{07743},
  \city{Jena},
  \cny{Germany}
}
\address[id=idea]{
  \orgname{Ideaconsult Ltd},
  \street{A. Kanchev 4},
  \postcode{1000},
  \city{Sofia},
  \cny{Bulgaria}
}
\address[id=leicester]{
  \orgname{Department of Informatics, University of Leicester},
  %\street{},
  %\postcode{LE1 7RH},
  \city{Leicester},
  \cny{UK}
}
\address[id=wi_mit]{
  \orgname{Whitehead Institute for Biomedical Research}, 
  \street{9 Cambridge Center},
  \postcode{MA 02142},
  \city{Cambridge},
  \cny{USA}
}
\address[id=miquel]{
  \orgname{Química Clínica Aplicada},
  %\street{4 Hanway Place},
  \postcode{43870},
  \city{Amposta},
  \cny{Spain}
}
\address[id=gilleain]{
  %\orgname{Department of Informatics, University of Leicester},
  \street{4 Hanway Place},
  \postcode{W1T 1HD},
  \city{London},
  \cny{UK}
}
\address[id=nih]{
  \orgname{National Center for Advancing Translational Science},
  \street{9800 Medical Center Drive},
  \postcode{MD 20878},
  \city{Rockville},
  \cny{USA}
}
\address[id=ebi]{
  \orgname{Metabolomics and Molecular Informatics, EMBL-EBI},
  %\street{A. Kanchev 4},
  %\postcode{1000},
  \city{Hinxton},
  \cny{UK}
}

@@

A chemical formula is the simplest chemical representation of a
compound.  It defines the number of isotopes or elements that
compose  a compound without describing how atoms
are bonded. With the rise of metabolomics it has become increasingly
relevant to have full support for these in cheminformatics
libraries~\cite{Wolf2010,RojasCherto2011,Pluskal2012,Pluskal2010}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}
The authors acknowledge the great number of people who have contributed smaller
and larger contributions to the CDK library. A full list of contributors is
found in the AUTHORS file~\cite{AUTHORS}. OS acknowledges support from the Swedish strategic research programs eSSENCE and Swedish e-Science Research Center (SeRC). TP is a Simons Foundation Fellow of the Helen Hay Whitney Foundation.

>>>>

\author[
   addressref={um},                                 % ORCID: 0000-0001-7542-0286
   %corref={aff1},
   %noteref={n1},
   email={egon.willighagen@maastrichtuniversity.nl}
]{\inits{ELW}\fnm{Egon L} \snm{Willighagen}}
\author[
   addressref={nm},
   email={john.may@cantab.net}
]{\inits{JWM}\fnm{John W} \snm{May}}
\author[
   addressref={uppsala},
   email={jonathan.alvarsson@farmbio.uu.se}
]{\inits{JA}\fnm{Jonathan} \snm{Alvarsson}}
\author[
   addressref={uppsala},
   email={berg.arvid@gmail.com}
]{\inits{AB}\fnm{Arvid} \snm{Berg}}
\author[
   addressref={azg},                                % ORCID: 0000-0001-9491-4134
   email={Lars.A.Carlsson@astrazeneca.com}
]{\inits{LC}\fnm{Lars}~\snm{Carlsson}}
% \author[
%    addressref={jena},
%    email={kai.duehrkop@uni-jena.de}
% ]{\inits{KD}\fnm{Kai} \snm{Dührkop}}
\author[
   addressref={idea},
   email={jeliazkova.nina@gmail.com}
]{\inits{NJ}\fnm{Nina} \snm{Jeliazkova}}
\author[
   addressref={leicester},
   email={shk12@le.ac.uk}
]{\inits{SK}\fnm{Stefan} \snm{Kuhn}}
\author[
   addressref={wi_mit},                             % ORCID: 0000-0002-6940-3006
   email={pluskal@wi.mit.edu}
]{\inits{TP}\fnm{Tomáš}~\snm{Pluskal}}
\author[
   addressref={miquel},                             % ORCID: 0000-0002-4659-1446
   email={mrojas@qca.es}
]{\inits{MRJ}\fnm{Miquel}~\snm{Rojas-Chertó}}
\author[
   addressref={uppsala},
   email={ola.spjuth@farmbio.uu.se}
]{\inits{OS}\fnm{Ola} \snm{Spjuth}}
\author[
   addressref={gilleain},                           % ORCID: 0000-0002-8368-6954
   email={gilleain.torrance@gmail.com}
]{\inits{GT}\fnm{Gilleain} \snm{Torrance}}
\author[
   addressref={um},
   email={chris.evelo@maastrichtuniversity.nl}
]{\inits{CTE}\fnm{Chris T} \snm{Evelo}}
\author[
   addressref={nih},
   email={guhar@mail.nih.gov}
]{\inits{RG}\fnm{Rajarshi} \snm{Guha}}
\author[
   addressref={ebi},
   email={steinbeck@ebi.ac.uk}
]{\inits{CS}\fnm{Christoph}~\snm{Steinbeck}}

\address[id=um]{
  \orgname{Dept of Bioinformatics - BiGCaT, NUTRIM, Maastricht University}, % university, etc
  %\street{Waterloo Road},
  \postcode{NL-6200 MD},
  \city{Maastricht},
  \cny{The Netherlands}
}
\address[id=nm]{
  \orgname{NextMove Software Ltd},
  %\street{D\"{u}sternbrooker Weg 20},
  \postcode{CB4 0EY},
  \city{Cambridge},
  \cny{UK}
}
\address[id=uppsala]{
  \orgname{Department of Pharmaceutical Biosciences, Uppsala University},
  %\street{D\"{u}sternbrooker Weg 20},
  \postcode{751 24},
  \city{Uppsala},
  \cny{Sweden}
}
\address[id=azg]{
  \orgname{AstraZeneca, Innovative Medicines \& Early Development, Quantitative Biology},
  %\street{},
  %\postcode{LE1 7RH},
  \city{Mölndal},
  \cny{SE}
}
% \address[id=jena]{
%   \orgname{Chair for Bioinformatics, Friedrich Schiller University},
%   %\street{},
%   \postcode{07743},
%   \city{Jena},
%   \cny{Germany}
% }
\address[id=idea]{
  \orgname{Ideaconsult Ltd},
  \street{A. Kanchev 4},
  \postcode{1000},
  \city{Sofia},
  \cny{Bulgaria}
}
\address[id=leicester]{
  \orgname{Department of Informatics, University of Leicester},
  %\street{},
  %\postcode{LE1 7RH},
  \city{Leicester},
  \cny{UK}
}
\address[id=wi_mit]{
  \orgname{Whitehead Institute for Biomedical Research}, 
  \street{9 Cambridge Center},
  \postcode{MA 02142},
  \city{Cambridge},
  \cny{USA}
}
\address[id=miquel]{
  \orgname{Química Clínica Aplicada},
  %\street{4 Hanway Place},
  \postcode{43870},
  \city{Amposta},
  \cny{Spain}
}
\address[id=gilleain]{
  %\orgname{Department of Informatics, University of Leicester},
  \street{4 Hanway Place},
  \postcode{W1T 1HD},
  \city{London},
  \cny{UK}
}
\address[id=nih]{
  \orgname{National Center for Advancing Translational Science},
  \street{9800 Medical Center Drive},
  \postcode{MD 20878},
  \city{Rockville},
  \cny{USA}
}
\address[id=ebi]{
  \orgname{Metabolomics and Molecular Informatics, EMBL-EBI},
  %\street{A. Kanchev 4},
  %\postcode{1000},
  \city{Hinxton},
  \cny{UK}
}

@@

A chemical formula is the simplest chemical representation of a
compound.  It defines the number of isotopes or elements that
compose  a compound without describing how atoms
are bonded. With the rise of metabolomics it has become increasingly
relevant to have full support for these in cheminformatics
libraries~\cite{Wolf2010,RojasCherto2011,Pluskal2012,Pluskal2010,Duhrkop2015}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}
The authors acknowledge the great number of people who have contributed smaller
and larger contributions to the CDK library. A full list of contributors is
found in the AUTHORS file~\cite{AUTHORS}. OS acknowledges support from the Swedish strategic research programs eSSENCE and Swedish e-Science Research Center (SeRC). TP is a Simons Foundation Fellow of the Helen Hay Whitney Foundation.
We also thank K. Dührkop for his contributions during the writing of this paper.
